### 2024-07-10 journal/related.tex
Removed repeated words
updating a bit of related work

%\Hossein{Maybe we can say worst-case. I see that you have used amortized step complexity consistently everywhere but maybe it is fair to say that these papers were not designed to improve the worst-case amortized step complexity.}
%\Eric{Amortized step complexity is defined in first parag of paper to be worst case over all executions of the average number of steps per operation, so I think the worst-case aspect of it is clear.  Also, if we just said ``worst-case'' without ``amortized'' it would be confused with worst-case of an individual operation, which is infinite in this case.  By the way, saying that previous implementations have $\Omega(p)$ step complexity, even in the amortized sense, makes the new contribution even more of a contrast.}

@@

Kogan and Herlihy \cite{DBLP:conf/podc/KoganH14} improved the MS-queue's performance using \emph{futures}.
Operations return future
objects instead of responses. Later, when an operation's response is needed, it
is evaluated using the future object.
This allows batches of enqueues or dequeues to be done at once on an MS-queue.
However, the implementation satisfies a weaker correctness condition than linearizability.
Milman-Sela et al.~\cite{MKLLP22} extended this approach to allow batches
to mix enqueues and dequeues.
%\here{Is this worth saying:
%They use some properties of the queue size before and after a batch, similar to a part of our work.}
In the worst case, where operations require their response right away,
batches have size 1, and both of these implementations behave like a standard MS-queue.

@@

\paragraph{Universal Constructions.}
\Eric{Rewrote this paragraph on Oct 25}
One can also build a queue using a universal construction \cite{10.1145/114005.102808}.
Afek, Dauber, and Touitou's universal construction~\cite{DBLP:conf/stoc/AfekDT95}
introduced the technique where a process must climb a binary tree while helping any other processes it sees along the way to reach the root,
and this technique is the basis of our queue implementation.
Jayanti \cite{DBLP:conf/podc/Jayanti98a} observed that
their construction can be modified to use $O(\log p)$ steps per operation, 
assuming that words can store $\Omega(p \log p)$ bits. 
However, if more reasonably-sized $O(\log p)$-bit words words are used, the construction would take $\Omega(p\log p)$ steps per operation.
There are two obstacles to making this construction more efficient:
it is expensive to manipulate lists of up to $p$ operations that must be propagated along a path
up to the root, and
when a batch of up to $p$ operations reaches the root, it is expensive to perform all of them on the
implemented data structure.
In our work, we devise a novel way of doing this implicitly for batches of operations on a queue.

@@

>>>>

@@

Kogan and Herlihy \cite{DBLP:conf/podc/KoganH14} improved the MS-queue's performance using \emph{futures}.
Operations return future
objects instead of responses. Later, when an operation's response is needed, it
is evaluated using the future object.
This allows batches of enqueues or dequeues to be done at once on an MS-queue.
However, the implementation satisfies a weaker correctness condition than linearizability.
Milman-Sela et al.~\cite{MKLLP22} extended this approach to allow batches
to mix enqueues and dequeues.
In the worst case, where operations require their response right away,
batches have size 1, and both of these implementations behave like a standard MS-queue.

@@

\paragraph{Universal Constructions.}
\Eric{Rewrote this paragraph on Oct 25}
One can also build a queue using a universal construction \cite{10.1145/114005.102808}.
Afek, Dauber, and Touitou's universal construction~\cite{DBLP:conf/stoc/AfekDT95}
introduced the technique where a process must climb a binary tree while helping any other processes it sees along the way to reach the root,
and this technique is the basis of our queue implementation.
Jayanti \cite{DBLP:conf/podc/Jayanti98a} observed that
their construction can be modified to use $O(\log p)$ steps per operation, 
assuming that words can store $\Omega(p \log p)$ bits. 
However, if more reasonably-sized $O(\log p)$-bit words are used, the construction would take $\Omega(p\log p)$ steps per operation.
There are two obstacles to making this construction more efficient:
it is expensive to manipulate lists of up to $p$ operations that must be propagated along a path
up to the root, and
when a batch of up to $p$ operations reaches the root, it is expensive to perform all of them on the
implemented data structure.
In our work, we devise a novel way of doing this implicitly for batches of operations on a queue.

@@

\HERE{discuss \cite{ACH18} who seem to have showed that strongly linearizable (define) queue is impossible using weak primitives; can also mention that HW22 observed that the MS queue is not strongly linearizable but can easily be made so}
